Network based multifactorial modelling microRNA-target interaction.

### 2019-10-08 NAR_form_article.Rmd
eps convertion of pics
article check revisions

\begin{figure*}[ht]
\begin{center}
\includegraphics[width=15cm,height=10cm]{Fig1.jpg}
\end{center}
\caption{Schematic presentation of mechanism of network based model. 
    (\textbf{A}) In steady state, miRNAs (M: triangles) repress targets (G: circles) according to proportion of target expression level. 
    (\textbf{B}) Two fold increase in transcript level of Gene2 (G2) acts as a trigger (shown in red). 
    (\textbf{C}) Distribution of miRNA1 (M1) changes. 
    (\textbf{D)}) The change at expression of common target affects changes of proportional distribution of miRNA2 (M2). Expression values are rounded to integers for simplicity. Gray values on edges indicate initial distribution, red values indicate increase and blue values indicate decrease in distribution. Shades of circles indicate different levels of increase.}
\label{fig:fig1}
\end{figure*}

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[width=9cm,height=17cm]{Fig2.jpg}
\end{center}
\caption{Calculations to determine of miRNA binding and repression efficiency. $G$, Gene; $M$, miRNA; $STE$, seed type effect; $RE$, Region Effect; $E$, Energy; $STE^\prime$, normalized values of seed type efficiency coefficient; $RE^\prime$, normalized values of region efficiency coefficient; $E^\prime$, normalized values of energy coefficient. Numbering on edges match the pair order in Table \ref{tab:one} }
\label{fig:fig2}
\end{figure}

@@

\begin{figure*}[ht]
\begin{center}
\includegraphics[width=15cm,height=8cm]{Fig3.jpg}
\end{center}
\caption{Target regulations with interaction parameters. 
     (\textbf{A}) In the steady-state the repression activity of miRNAs on the targets after binding and repression efficiency. 
     (\textbf{B}) The changes the repression activities after increasing of G2 expression. 
     (\textbf{C}) Perturbation of primary neighborhoods of M1 miRNA (M1 miRNA group). 
     (\textbf{D}) Regulation of gene expression of other gene group via triggering target (common target between M1 and M2)}
\label{fig:fig3}
\end{figure*}

## ceRNA:Target networks based on interaction factors
Earlier studies reported that miRNA regulatory interactions are affected by different parameters. For example, Xu et al. have investigated the importance of seed pairing type between miRNAs and their targets and target site location by using proteomics dataset [@xu_characterization_2014]. They have proposed that the features of binding between miRNA and target can be critical for miRNA efficiency. In addition, binding energy between miRNA and targets is a significant determinant for miRNA efficiency and it has been reported that strength of miRNA:target interactions is depended on binding energy of complexes [@breda_quantifying_2015]. Similarly, another study revealed that affinity is correlated with seed pairing of miRNA:target pairs and suggested that affinity is correlated with number of canonical seed base pairing [@bosson_endogenous_2014]. Therefore, we have integrated aforementioned interaction parameters that could be useful for predicting miRNA repression activity more accurately. The sample dataset used in Figure \ref{fig:fig1} is recalculated with additional factors Table \ref{tab:one} in effect with same trigger (Detailed codes can be found at Section 3 in Supplementary Materials and Methods), two fold increase in Gene2 (Figure \ref{fig:fig3}B, and Figure S6). When the factors were taken into account in the system, miRNA efficiencies varied as shown in Figure \ref{fig:fig3}A (Figure S5). Although the miRNA:target expression ratios in steady-state were same in comparison with the sample dataset without factors, efficiency of binding and repression have changed.  On the other hand, changes in expression levels of G4 (Detailed codes are available at Section 4 in Supplementary Materials and Methods; and Figure S8) and G1 differ although they have same initial conditions. This is due to fact that G4 is targeted by two miRNAs, M1 and M2 (Figures \ref{fig:fig3}D) and changed interaction factors.

@@

We have analysed the realistic network in context of perturbation efficiencies of nodes. This process were run with 3 fold upregulation and 10 iteration in conditions where elements that have less than five percent change are not taken into account (Codes are found at Section 6.5 in Supplementary Materials and methods). With this approach, SERPINE2 has been found as one of the most effective gene nodes in network. We have observed that gaining steady-state in realistic network takes a lot of time because of complex interactions, as expected, and SERPINE2 exhibit highest efficiency at iteration 31 with given conditions (see Section 6.6 in Supplementary Materials and Methos and Figure S9 in Supplementary Figures). We ran the model with two fold of iteration which is observed as most effective. We have observed that the results of simulation with SERPINE2 gene, a serine protease inhibitor, one of the overexpressed genes of breast cancer [@yang2018expression; @candia2006protease] is not convenient with tumor tissue expression values, when we compared the tumor gene expression dataset retrieved from TCGA and simulation results obtained via our approach (see Section 6.6 in Supplementary Materials and methods). Simulation results were not expected to be consistent with tumor tissue expression values because large number of abnormal regulations were observed in these tissues. So, it might not said that a single gene is responsible for entire regulations in cancer cells. Besides, the all other factors such as up/down regulation of miRNAs or other genes were ignored. Therefore, it would be more useful to test our approach with an ideal dataset that includes entire informations of gene expression values before and after the regulation of a known gene at existing tissue, and miRNA expression values of this tissue at initial conditions.

>>>>

\begin{figure*}[ht]
\begin{center}
\includegraphics[width=16cm,height=9cm]{Fig1.eps}
\end{center}
\caption{Schematic presentation of mechanism of network based model. 
    (\textbf{A}) In steady state, miRNAs (M: triangles) repress targets (G: circles) according to proportion of target expression level. 
    (\textbf{B}) Two fold increase in transcript level of Gene2 (G2) acts as a trigger (shown in red). 
    (\textbf{C}) Distribution of miRNA1 (M1) changes. 
    (\textbf{D)}) The change at expression of common target affects changes of proportional distribution of miRNA2 (M2). Expression values are rounded to integers for simplicity. Gray values on edges indicate initial distribution, red values indicate increase and blue values indicate decrease in distribution. Shades of circles indicate different levels of increase.}
\label{fig:fig1}
\end{figure*}

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[width=9cm,height=17cm]{Fig2.eps}
\end{center}
\caption{Calculations to determine of miRNA binding and repression efficiency. $G$, Gene; $M$, miRNA; $STE$, seed type effect; $RE$, Region Effect; $E$, Energy; $STE^\prime$, normalized values of seed type efficiency coefficient; $RE^\prime$, normalized values of region efficiency coefficient; $E^\prime$, normalized values of energy coefficient. Numbering on edges match the pair order in Table \ref{tab:one} }
\label{fig:fig2}
\end{figure}

@@

\begin{figure*}[ht]
\begin{center}
\includegraphics[width=15cm,height=8cm]{Fig3.eps}
\end{center}
\caption{Target regulations with interaction parameters. 
     (\textbf{A}) In the steady-state the repression activity of miRNAs on the targets after binding and repression efficiency. 
     (\textbf{B}) The changes the repression activities after increasing of G2 expression. 
     (\textbf{C}) Perturbation of primary neighborhoods of M1 miRNA (M1 miRNA group). 
     (\textbf{D}) Regulation of gene expression of other gene group via triggering target (common target between M1 and M2)}
\label{fig:fig3}
\end{figure*}

## ceRNA:Target networks based on interaction factors
Earlier studies reported that miRNA regulatory interactions are affected by different parameters. For example, Xu et al. have investigated the importance of seed pairing type between miRNAs and their targets and target site location by using proteomics dataset [@xu_characterization_2014]. They have proposed that the features of binding between miRNA and target can be critical for miRNA efficiency. In addition, binding energy between miRNA and targets is a significant determinant for miRNA efficiency and it has been reported that strength of miRNA:target interactions is depended on binding energy of complexes [@breda_quantifying_2015]. Similarly, another study revealed that affinity is correlated with seed pairing of miRNA:target pairs and suggested that affinity is correlated with number of canonical seed base pairing [@bosson_endogenous_2014]. Therefore, we have integrated aforementioned interaction parameters that could be useful for predicting miRNA repression activity more accurately. The sample dataset used in Figure \ref{fig:fig1} is recalculated with additional factors Table \ref{tab:one} in effect with same trigger (Detailed codes can be found at Section 3 in Supplementary Materials and Methods), two fold increase in Gene2 (Figure \ref{fig:fig3}B, and Figure S6). When the factors were taken into account in the system, miRNA efficiencies varied as shown in Figure \ref{fig:fig3}A (Figure S5). Although the miRNA:target expression ratios in steady-state were same in comparison with the sample dataset without factors, efficiency of binding and repression have changed.  On the other hand, changes in expression levels of G4 (Detailed codes are available in Figure S8 and Section 4 in Supplementary Materials and Methods) and G1 differ although they have same initial conditions. This is due to fact that G4 is targeted by two miRNAs, M1 and M2 (Figure \ref{fig:fig3}D) and changed interaction factors.

@@

We have analysed the realistic network in context of perturbation efficiencies of nodes. This process was run with 3 fold upregulation and 10 iteration in conditions where elements that have less than five percent change are not taken into account (Codes are found at Section 6.5 in Supplementary Materials and Methods). With this approach, SERPINE2 has been found as one of the most effective gene nodes in network. We have observed that gaining steady-state in realistic network takes a lot of time because of complex interactions, as expected, and SERPINE2 exhibit highest efficiency at iteration 31 with given conditions (see Section 6.6 in Supplementary Materials and Methos and Figure S9 in Supplementary Figures). We ran the model with two fold of iteration which is observed as most effective. We have observed that the results of simulation with SERPINE2 gene, a serine protease inhibitor, one of the overexpressed genes of breast cancer [@yang2018expression; @candia2006protease] is not convenient with tumor tissue expression values, when we compared the tumor gene expression dataset retrieved from TCGA and simulation results obtained via our approach (see Section 6.6 in Supplementary Materials and Methods). Simulation results were not expected to be consistent with tumor tissue expression values because large number of abnormal regulations were observed in these tissues. So, it might not said that a single gene is responsible for entire regulations in cancer cells. Besides, the all other factors such as up/down regulation of miRNAs or other genes were ignored. Therefore, it would be more useful to test our approach with an ideal dataset that includes entire informations of gene expression values before and after the regulation of a known gene at existing tissue, and miRNA expression values of this tissue at initial conditions.
